---
title       : Random Geospatial Data Generator
subtitle    : A small (and unfinished) Shiny R App as the Coursera Assignment
author      : Ingo Nader
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [mathjax]            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
---

```{r r-pre, echo=FALSE, results="hide", warning=FALSE, message=FALSE}
## +++++++++++++++++++++++++++++++++++++
## run before knitting:
## +++++++++++++++++++++++++++++++++++++
path.slides <- "C:/data-sync/coursera/data-science-09-data-products/assignment01/pres-slidify2"
#setwd(path.slides)
library(slidify)
## +++++++++++++++++++++++++++++++++++++

library(XML)
library(ggplot2)
library(ggmap)
library(mapproj)
library(XML)
library(gdata)

## fake input object:
input <- list("sel.country"="Deutschland", sel.nrows=5000)

deg.to.dec <- function(instr) {
  instr <- gsub(pattern="[Â]", replacement="", x=instr)
  #instr <- gsub(pattern=" \\(", replacement="(", x=instr)
  instr <- trim(instr)
  instr.split <- strsplit(instr, split="[°´'`]")
  instr.deg <- as.numeric(sapply(instr.split, function(i) i[1]))
  instr.min <- as.numeric(sapply(instr.split, function(i) i[2]))
  instr.sign <- sapply(instr.split, function(i) ifelse(trim(i[3]) %in% c("N", "O"), 1, -1))
  ret <- instr.sign * (instr.deg + instr.min/60)
  return(ret)
}
#deg.to.dec("44° 24´ N")
#deg.to.dec("50Â° 44Â´ N")

adj.range <- function(x, adj=1.1) {
  r <- range(x)
  r.adj <- (r - mean(r)) * adj + r
}


## ============================================================================
## read list of geolocations
## ============================================================================

doc.text <- readLines("geolocs-astrologon.html", encoding="ISO-8859-1")
doc <- htmlParse(doc.text)

## see the whole node, including html tags:
tab.nodes.lines <- getNodeSet(doc, "//table/tr") 

tab.nodes.lines.entries <- lapply(tab.nodes.lines[-1], getNodeSet, "td")
#tab.nodes.lines.entries[1:2]

loc.raw <- sapply(tab.nodes.lines.entries, function(i) trim(xmlValue(i[[1]])))
lat.raw <- sapply(tab.nodes.lines.entries, function(i) trim(xmlValue(i[[2]])))
lon.raw <- sapply(tab.nodes.lines.entries, function(i) trim(xmlValue(i[[3]])))

locbase <- data.frame(
  "city"=sapply(trim(strsplit(loc.raw, ",")), function(i) i[1]),
  "country"=sapply(trim(strsplit(loc.raw, ",")), function(i) i[2]),
  "lat"=deg.to.dec(lat.raw),
  "lon"=deg.to.dec(lon.raw))

    ## select country:
    locbase.sel <- locbase[grep(input$sel.country, locbase$country, fixed=TRUE),]
    
    ## generate a sample of data points for the selected country:
    loc.idx <- sample(seq_along(locbase.sel$city), size=input$sel.nrows, replace=TRUE)
    
    ## construct the data frame
    dat.all <- data.frame(idx = seq_along(loc.idx))
    dat.all$country <- locbase.sel$country[loc.idx]
    dat.all$city <- locbase.sel$city[loc.idx]
    dat.all$lat <- (locbase.sel$lat[loc.idx] + rnorm(input$sel.nrows, mean=0, sd=0.25))
    dat.all$lon <- (locbase.sel$lon[loc.idx] + rnorm(input$sel.nrows, mean=0, sd=0.25))

    rng.lat <- adj.range(dat.all$lat, adj=1.2)
    rng.lon <- adj.range(dat.all$lon, adj=1.2)
    bbox <- c(min(rng.lon), min(rng.lat), max(rng.lon), max(rng.lat))
    

```

## What the App does

**Main Purpose** of this app:
* generates random points around some cities
* to be used for visualization purposes if you don't have data

**How it's done** (overview):
* Step 1: get the GPS coordinates 
  * from a `.html` file (included with the app, 
but downloaded from [here](http://www.google.com.tr/url?sa=t&rct=j&q=&esrc=s&source=web&cd=1&cad=rja&uact=8&ved=0CBwQFjAA&url=http%3A%2F%2Fwww.astrologon.com%2Fpositionen.htm&ei=5PufU-WSH-fy7AaWm4CwDQ&usg=AFQjCNHS_L79tP4JByIyzX7WvvTAUV19bg&bvm=bv.68911936,d.ZGU))
  * and parse the file to actually get the coordinates in lat/lon format
* Step 2: Generate points by adding random error to lat/lon coordinates
* Step 3: download the corresponding part from Google Maps
* Step 4: Display

--- 

## Step 1: Get GPS Coordinates

* using `XML` package's `htmlParse()` function to get the nodes
* then decoding it using string functions (omitted for brevity)

```{r get-gps}
doc.text <- readLines("geolocs-astrologon.html", encoding="ISO-8859-1")
doc <- htmlParse(doc.text)
tab.nodes.lines <- getNodeSet(doc, "//table/tr")  ## see the whole node, including html tags:
tab.nodes.lines.entries <- lapply(tab.nodes.lines[-1], getNodeSet, "td")
tab.nodes.lines.entries[[1]]
```

--- 

## Step 2: Generate Points

In a `reactive({})` function:
* using the `input` object, select the data of the selected country
* generate a sample of `nrows` objects
* construct data frame with lat/lon objects plus $N(0,1)$ random error

```{r gen-points, eval=FALSE}
construct.data <- reactive({
  locbase.sel <- locbase[grep(input$sel.country, locbase$country, fixed=TRUE),]
  loc.idx <- sample(seq_along(locbase.sel$city), size=input$sel.nrows, replace=TRUE)
  dat.all <- data.frame(idx = seq_along(loc.idx))
  dat.all$country <- locbase.sel$country[loc.idx]
  dat.all$city <- locbase.sel$city[loc.idx]
  dat.all$lat <- (locbase.sel$lat[loc.idx] + rnorm(input$sel.nrows, mean=0, sd=0.25))
  dat.all$lon <- (locbase.sel$lon[loc.idx] + rnorm(input$sel.nrows, mean=0, sd=0.25))
  return(dat.all)   
})
```


--- 

## Step 3: Download from Google Maps

In a `renderPlot({})` function:
* calculate a bounding box `bbox` for the geospatial data
* and construct the plot object (**NOTE**: This may take some time...)

```{r map, cache=TRUE, warning=FALSE}
    map.google <- get_map(location = bbox, zoom = 6)

    p2 <- ggmap(map.google) + labs(title = input$wch.country)
    p2 <- p2 + geom_point(data=dat.all, aes(y = lat, x = lon), alpha=0.2, size=2, col="red")
 
```
--- 

## Step 4: Display

```{r display, cache=TRUE}
    plot(p2)
```


